\paragraph{A specification is only as strong as its keywords.}
The most prevalent divergence we measure---unknown-tool calls returned as
\texttt{isError} results rather than JSON-RPC protocol errors---is not, strictly, a
violation, because the specification never says \textsf{MUST}. It categorizes the
behavior in prose and shows one example. That single editorial choice, multiplied
across the official SDKs that had to pick an interpretation, produced an ecosystem
in which the written categorization is descriptively false for the large majority of
servers. The lesson generalizes beyond MCP: in a protocol whose conformance is
mediated by a few dominant SDKs, non-normative language does not leave behavior
unconstrained---it delegates the constraint to the SDK authors. Where the
specification wants uniformity, it must say \textsf{MUST}.

\paragraph{The SDK is the ecosystem.}
Attributing servers to their SDK collapses thousands of independent authorship
decisions into a handful of library defaults. This is good news and bad news. The
bad news is that a single suboptimal default propagates to everyone. The good news
is that the ecosystem is correspondingly \emph{fixable at the source}: a change to a
handful of SDKs, or a clarified specification plus updated SDK defaults, would move
the conformance numbers for the whole population at once. Our results argue for
treating SDK defaults as ecosystem policy and testing them against the spec directly.

\paragraph{Startup failure is an availability tax.}
That a large fraction of installable servers never reach a handshake reframes
``server selection'' for agent builders as an empirical reliability problem, not a
catalog lookup. Registries rank servers by popularity and recency; none, today,
surface whether a server \emph{starts}. A pre-publish conformance gate---exactly the
kind the community has begun to discuss---would convert our one-time measurement into
a standing quality signal.

\paragraph{Implications.}
For SDK authors: revisit the unknown-tool and argument-validation defaults, and
consider emitting protocol errors where the spec's own categorization points.
For the specification: promote the error-handling categorization to normative
language, or explicitly bless both mechanisms. For registries: add a conformance and
liveness signal. For agent frameworks: treat wrong-typed-argument acceptance and
malformed-input fragility as client-side risks to defend against, since a
non-trivial minority of servers exhibit them.
