This study executes third-party code and probes servers for defects, so we adopt
the norms of responsible measurement research.

\paragraph{Safe execution.}
Every server---during both installation and probing---runs inside an isolated
container with all Linux capabilities dropped, no new privileges, bounded memory,
CPU, and process counts, and no host filesystem access beyond a package cache
volume. The install phase, which executes arbitrary \texttt{postinstall} and build
scripts with network access, receives the same isolation as the probe phase; the
probe phase additionally runs with networking disabled. We execute no server that
declares a requirement for credentials, both because it lies outside our frame and
because doing so could send real requests to third-party systems.

\paragraph{Responsible disclosure.}
A subset of our findings are security-relevant: a server that crashes or hangs on a
malformed frame is a denial-of-service primitive, and a server that emits
non-protocol bytes on its stdio channel can desynchronize a client. We report these
only in aggregate and through anonymized case studies. We do not publish a
name-and-shame list of individual servers. Security-relevant findings are recorded
in a private disclosure log with maintainer contacts resolved from registry
repository metadata, and are reported to maintainers before publication; where a
finding is exploitable, per-server details are withheld until a reasonable
disclosure window has elapsed. The released dataset is filtered accordingly.

\paragraph{Registry and maintainer impact.}
Our crawl uses the registry's public API within its rate limits, and our installs
draw from public package registries in the ordinary way. We cache aggressively to
avoid redundant load. We view the released harness as a benefit to maintainers: it
lets any author check their own server's conformance before publishing, which is
the outcome the community's proposed pre-publish conformance checklist seeks.
