Large language model agents increasingly act on the world through external tools,
and the Model Context Protocol (MCP) has become the dominant standard for
connecting them. In under two years MCP has gone from a single vendor's proposal
to an ecosystem with public registries listing tens of thousands of servers; the
official registry alone contains \Nframe{} unique servers at the time of our study.
When an agent selects a tool and an MCP server mishandles the request---returns a
malformed result, accepts ill-typed arguments, or simply fails to start---the
agent does not merely lose a capability; it may proceed on wrong output with full
confidence. The reliability of this server layer is therefore a reliability
property of every agent built on top of it.

Yet we know remarkably little about how these servers actually behave, because
every prior empirical study of the MCP ecosystem examines servers \emph{without
running them}. Existing work scans source code and metadata for health and
maintainability, crawls registries to characterize scale and supply-chain
structure, or mines GitHub issues to taxonomize faults reported by users. These
static lenses answer valuable questions, but they cannot answer a prior and
strikingly basic one: \emph{if you install these servers and speak the protocol to
them, do they work, and do they follow the specification?} The protocol maintainers
ship a conformance test suite, but it exercises the handful of official SDKs in
continuous integration; it has never been pointed at the deployed ecosystem. Put
plainly: the official suite tests the five SDKs, and no one tests the seventeen
thousand servers built on them.

We close that gap. We present the first execution-based conformance study of the
MCP server ecosystem. Starting from the official registry, we draw a random sample
of self-contained, locally-installable servers, launch each in a network-isolated
sandbox, and drive it through a conformance harness that exercises the
initialization lifecycle, tool listing, declared-schema validity, and behavior
under three inputs that arise in ordinary agent sessions: a call to an unknown
tool, a call with a wrong-typed argument, and a malformed protocol frame. We grade
each server against the protocol version it negotiates, and we take care to
separate genuine specification requirements (RFC-2119 \textsf{MUST}/\textsf{SHOULD})
from behavior the specification merely illustrates by example.

Our findings are consistent and, in places, striking. Only \HandshakeRate{} of
installable servers successfully start and complete a protocol handshake; the
remainder fail in ways we classify into a startup-failure taxonomy, from install
errors to silent exits to hangs. Among servers that do respond, \ErrAsResultRate{}
report a call to a non-existent tool as a ``successful'' tool-execution error
(\texttt{isError: true}) rather than the JSON-RPC protocol error the specification
illustrates---and we trace this near-universal divergence to its cause: it is the
default behavior of the official SDKs, inherited wholesale by the servers built on
them. A smaller but important minority silently accept ill-typed arguments or fail
on malformed input, the latter a denial-of-service primitive. Throughout, we find
that the gap between the written specification and deployed practice is wide,
systematic, and explicable.

This paper makes four contributions:
\begin{itemize}
  \item \textbf{The first execution-based, protocol-level conformance measurement of
    the MCP ecosystem} (C1), at a scale and level of reproducibility not previously
    attempted, releasing an open harness and dataset.
  \item \textbf{A two-dimensional taxonomy} (C2) of startup-failure classes and
    conformance-violation classes, each tied to a concrete agent-facing consequence.
  \item \textbf{A root-cause account of the spec--practice divergence} (C3): by
    attributing each server to its SDK, we show the dominant divergence is
    institutionalized in the official SDKs rather than independently reinvented.
  \item \textbf{Open infrastructure} (C4): a conformance harness that extends the
    official suite's scenarios to arbitrary deployed servers, plus the released
    measurement dataset, enabling the ecosystem to monitor its own conformance.
\end{itemize}
