We measure the MCP server ecosystem by \emph{executing} it. This section defines
the sampling frame, the conformance harness, the sandbox, and the grading rules.
The harness and the released dataset make every number reproducible.

\subsection{Sampling frame}
We crawl the official MCP registry (\texttt{registry.modelcontextprotocol.io})
via its cursor-paginated API, retaining the latest published version of each
server. This yields \Nframe{} unique servers. Each server declares zero or more
\emph{packages} (installable artifacts on npm, PyPI, OCI, NuGet, or Cargo) and/or
\emph{remotes} (hosted HTTP endpoints). Our study population is the set of servers
that are (i) locally installable, (ii) use the \texttt{stdio} transport, (iii)
distributed on npm or PyPI (the two dominant registries), and (iv)
\emph{self-contained}: they declare no required environment variable or argument
that lacks a default. Criterion (iv) restricts us to servers that a client could
launch without provisioning secrets, which both bounds the ethics of mass
execution and defines a clean, reproducible population. We report the exclusion of
remote-only and credential-requiring servers explicitly (Section~\ref{sec:threats})
and treat the resulting frame as the denominator for all rates. From this frame we
draw a fixed pseudo-random sample (seeded) of \Nprobed{} servers.

\subsection{Conformance harness}
The harness, \texttt{mcpprobe}, is a deliberately minimal, hand-written JSON-RPC
client speaking line-delimited messages over the server's stdio. We do \emph{not}
use an official SDK client: to observe raw behavior---including responses to
malformed frames---the probe must not inherit any SDK-side correction or
validation. The harness offers protocol version \texttt{2025-06-18} at
initialization, records the version the server \emph{negotiates}, and grades all
subsequent checks against that negotiated version rather than the latest
specification.

Each server is driven through the following checks, whose identifiers align with
the categories of the official \texttt{modelcontextprotocol/conformance} suite
where an analogue exists:

\begin{itemize}
  \item \textbf{server-initialize}: the \texttt{initialize} request must return a
    result carrying a protocol version and a capabilities object; the harness then
    sends \texttt{notifications/initialized}.
  \item \textbf{ping}: a \texttt{ping} must return an empty-object result.
  \item \textbf{tools-list}: if the server advertises the tools capability,
    \texttt{tools/list} must return an array in which every entry has a
    \texttt{name} and an \texttt{inputSchema} object.
  \item \textbf{tools-schema-valid}: each declared \texttt{inputSchema} must be a
    valid JSON Schema (checked by meta-validation).
  \item \textbf{tools-call-unknown}: calling a non-existent tool must be rejected.
    The specification's Error Handling section categorizes ``Unknown tools'' under
    \emph{Protocol Errors} (standard JSON-RPC errors) and illustrates this with a
    \texttt{-32602} example, as opposed to \emph{Tool Execution Errors} reported as
    a result with \texttt{isError: true}. Crucially, this categorization is prose
    plus an example; it uses no RFC-2119 \textsf{MUST}/\textsf{SHOULD} keyword. We
    therefore do not treat a non-conforming response as a ``violation''; instead we
    record which of the two mechanisms the server uses (see
    \textsf{error-as-result} below).
  \item \textbf{tools-call-invalid-args}: for a tool with a typed required
    property, calling it with a wrong-typed value should be rejected. The spec is
    itself ambiguous here---it lists ``Invalid arguments'' under Protocol Errors but
    ``Invalid input data'' under Tool Execution Errors---so we count \emph{either} a
    protocol error \emph{or} an \texttt{isError} result as acceptable, and flag only
    a plain success response (the tool ran on ill-typed input) as a defect.
  \item \textbf{malformed-json}: after receiving a syntactically invalid frame, the
    server must not crash or hang; a subsequent \texttt{ping} must still succeed.
  \item \textbf{stdout-purity}: under the stdio transport, the server must emit only
    protocol messages on stdout; any non-JSON line is a channel-corruption defect.
\end{itemize}

\subsection{Grading and the error-as-result category}
Verdicts are \textsf{pass}, \textsf{fail}, \textsf{warn}, \textsf{skip}, and one
category introduced by calibration: \textsf{error-as-result}. When validating the
harness against the official reference servers (\texttt{server-memory},
\texttt{server-everything}), we found they answer an unknown-tool call not with the
specified JSON-RPC \texttt{-32602} error but with a \emph{successful} response whose
\texttt{result.isError} flag is set. Because this behavior originates in the
official SDKs and is therefore ecosystem-wide, we grade it as its own category
rather than a failure, and quantify its prevalence as a first-class result: it is
the clearest instance of specification text diverging from universal practice.

\subsection{Sandbox}
Executing thousands of untrusted packages demands isolation. Each server runs in a
fresh container (\texttt{node:22-slim} for npm, an \texttt{astral-sh/uv} image for
PyPI) with a memory cap, a single CPU, a PID limit, \texttt{no-new-privileges}, and
no host filesystem mounts other than a package-cache volume. Measurement uses a
\textbf{two-phase} protocol: an online install pass populates the cache, then the
conformance probe runs with \texttt{-{}-network=none}. No server code has network
access while it is being measured, which both hardens the experiment and removes
network flakiness and server-side outbound calls as confounders.

\subsection{Funnel accounting}
Because ``does it even run'' is itself a research question, we count every stage:
declared $\rightarrow$ eligible (self-contained stdio npm/PyPI) $\rightarrow$
sampled $\rightarrow$ container started $\rightarrow$ process launched
$\rightarrow$ handshake completed $\rightarrow$ probed. Servers that never reach a
handshake are classified into startup-failure categories
(Section~\ref{sec:results}) from their exit code and stderr, so that the yield loss
is characterized rather than merely reported.
