\subsection{The Model Context Protocol}
MCP standardizes how LLM agents discover and invoke external tools. A client and
server negotiate a protocol version and capabilities during an
\texttt{initialize} handshake, after which the client can list tools
(\texttt{tools/list}) and invoke them (\texttt{tools/call}). Messages follow
JSON-RPC 2.0, including its error-object semantics and reserved error codes. The
most common deployment is a local server communicating over stdio, where stdout
carries protocol traffic exclusively. The specification is versioned by date, and
servers may negotiate down to a version they support.

\subsection{Static studies of the MCP ecosystem}
The ecosystem has been studied at scale, but statically. Health-and-maintainability
work analyzes source and metadata of roughly 1,900 servers with static scanners
and code-smell detectors~\cite{mcpfirstglance}. Registry-measurement work crawls
multiple marketplaces and characterizes scale, maintenance, and supply-chain
concentration for on the order of 8,000 servers, but from listing metadata rather
than execution~\cite{mcpcrawler}. Fault taxonomies derive categories of runtime
misbehavior by \emph{mining GitHub issues}~\cite{runtimefaults,realfaults}; these
report that tool faults and schema-enforcement problems dominate the fault
landscape---smoke that our harness measures directly as fire. None of these works
executes servers to test conformance.

\subsection{Execution-based work is security-focused}
Where prior work does execute MCP servers, it targets security rather than
conformance, and on small or synthetic populations. Sandboxed behavioral scanners
and taint-style analyzers hunt vulnerabilities on curated or intentionally
vulnerable server sets~\cite{sandboxscan,viper}; a security-invariants benchmark
\emph{deliberately avoids live third-party execution}, using fixtures and a
40-repository metadata sample~\cite{securityinvariants}; malicious-server detection
and authentication studies likewise scope to security surfaces~\cite{maliciousmcp,
remoteauth}. Protocol conformance---whether a server that is neither malicious nor
vulnerable actually behaves as the specification requires---is a distinct question
that this line of work does not address.

\subsection{Conformance tooling and the gap we fill}
The protocol maintainers ship an official conformance suite, but it tests the
handful of official SDKs in continuous integration; it has not been pointed at the
deployed ecosystem. The community has begun proposing a pre-publish conformance
checklist for server authors, evidencing demand for exactly the measurement we
provide. Practitioner audits have reported striking headline numbers (e.g., that a
majority of listed servers are unmaintained), but without published methodology,
data, or protocol-level testing. To our knowledge, ours is the first reproducible,
protocol-level, execution-based conformance measurement of the MCP server
ecosystem, released with an open harness and dataset. In spirit it follows a long
line of empirical software-engineering measurement studies that execute a
population to characterize how it really behaves, rather than how its documentation
says it should.
