\label{sec:threats}

\paragraph{Construct validity.}
Our verdicts operationalize ``conformance,'' and a poorly chosen operationalization
could over- or under-count. We mitigate this in three ways. First, we grade each
server against the protocol version it \emph{negotiates}, not the latest
specification, so a server is never penalized for a requirement that postdates its
target revision. Second, we distinguish normative requirements from illustrative
ones: the most prevalent divergence we report (unknown-tool responses) is graded as
a distinct \textsf{error-as-result} category rather than a failure, precisely
because the specification expresses it in prose and an example rather than an
RFC-2119 keyword. Third, we calibrated the harness against the official reference
servers before the study and confirmed it reproduces their known-good behavior on
every check except the SDK-level divergence we set out to measure.

\paragraph{Internal validity.}
Measurement runs under network isolation (\texttt{-{}-network=none}), which removes
network flakiness and server-side outbound calls as confounders but could, in
principle, penalize a server that requires network access during initialization.
Because our frame excludes servers declaring required credentials or arguments, and
because the checks we grade (handshake, tool listing, error handling) are local
protocol behaviors, we expect this effect to be small; servers whose behavior
demonstrably requires network access are recorded as such rather than silently
scored as failures. Sandbox heterogeneity across language runtimes is a further
source of noise, which we bound by pinning base images and package versions and by
stamping every result with the harness commit that produced it.

\paragraph{External validity.}
Our population is self-contained, locally-installable servers distributed on npm and
PyPI using the stdio transport. This excludes remote-only servers (which we cannot
execute without provisioning hosted infrastructure) and servers requiring
credentials (which we exclude for both ethics and reproducibility). We report the
size of each excluded class explicitly and confine all rates to the stated frame;
we do not claim they generalize to remote or credentialed servers, whose conformance
is important future work. Because agent clients overwhelmingly launch local stdio
servers, we argue this population is the most consequential one to characterize
first.

\paragraph{Sample size and scope.}
We analyze a random sample of \Nprobed{} servers (\Nresponders{} responding) rather
than the full eligible frame. This sample sizes our confidence intervals (reported
throughout) but is a first execution-based measurement, not a census; the harness is
released precisely so a larger uniform-condition run---which a modest cloud host can
complete in hours---can tighten every estimate, most importantly the absolute
install-failure rate in the RQ1 funnel. We regard the conformance findings (RQ2--RQ4)
as robust---they replicate independently across our two probe conditions and are
near-universal---and the runnability magnitude (RQ1) as a well-motivated preliminary
estimate.

\paragraph{Excluded runs.}
A subset of probe attempts was invalidated by transient failures of the sandbox host
(the container engine), which we detect by an engine-error signature in the launch
transcript that is distinct from any package or server error. These carry no
information about the server under test and are excluded before analysis; because the
exclusion criterion is a property of the harness environment and not of the server,
it does not bias the reported rates. The released dataset flags them for audit.

\paragraph{Reproducibility.}
Every number in this paper is regenerable from released raw JSON-RPC transcripts and
the pinned harness commit. The sampling seed, registry snapshot, base images, and
package versions are all recorded. Because the registry is a moving target, we
report the crawl date and treat the snapshot, not the live registry, as the object
of study.
